---
title: "data visualization (smoothing qplot)"
author: "Pradip Basnet"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

#Add a Smoother

```{r}
data("diamonds")
```


tas 1:create a smaller dataset from diaminds by randomly sample. create a reproducible random sample of 100 rows from the diamonds dataset
```{r}
set.seed(1000)
dsmall <- diamonds[sample(nrow(diamonds),100),]
```


task 2:visualize the relationship between the logarithmically transformed carat weight (log(carat)) and logarithmically transformed price (log(price)) of diamonds in the dsmall dataset using a scatter plot with a smoothed trend line
```{r}
qplot(log(carat),log(price),data=dsmall,geom = "smooth")
```
task 3:combine multiple geoms using the c() function, visualize the relationship between the logarithmically transformed carat weight (log(carat)) and logarithmically transformed price (log(price)) of diamonds in the dsmall dataset using a scatter plot with individual points and a smoothed trend line
```{r}
qplot(log(carat),log(price),data = dsmall,geom = c("point","smooth"))
```
task 4:visualize the relationship between the logarithmically transformed carat weight (log(carat)) and logarithmically transformed price (log(price)) of diamonds in the dsmall dataset using a scatter plot with individual points and a more sensitive smoothed trend line. 
```{r}
qplot(log(carat),log(price),data = dsmall,geom = c("smooth","point"),span=0.2)
```
task 5:Setting span = 1 in this context will create a very smooth trend line that reflects the general relationship between log(carat) and log(price) in the dsmall dataset
```{r}
qplot(log(carat),log(price),data = dsmall,geom = c("smooth","point"),span=1)
```
installing and loading the necessary package
```{r}
#install.packages("splines")
library("splines")
```

##fit a linear model instead by using method = “lm” or robust linear model “rlm” which is less sensitive to outlier

task 6:visualize the relationship between the logarithmically transformed carat weight (log(carat)) and logarithmically transformed price (log(price)) of diamonds in the dsmall dataset using a scatter plot with individual points and a linear regression trend line
```{r}
qplot(log(carat), log(price), data = dsmall, geom = c("point", "smooth"), method = "lm")
```
task 7: visualize the relationship between engine displacement (displ) and highway miles per gallon (hwy) for cars in the mpg dataset using scatterplot
```{r}
qplot(displ,hwy,data = mpg)
```

task 7: generate a scatter plot where the color of each point is determined by the drv variable
```{r}
qplot(displ, hwy, data = mpg, colour = drv)
```

task 8:visualize the relationship between engine displacement (displ) and highway miles per gallon (hwy) for cars in the mpg dataset using a scatter plot with individual points and a smoothed trend line
```{r}
qplot(displ, hwy, data = mpg, geom = c("point", "smooth"))
```
##How are engine size and fuel economy related?

task 9:visualize the relationship between engine displacement (displ) and highway miles per gallon (hwy) for cars in the mpg dataset, with points colored according to the number of cylinders
```{r}
qplot(displ,hwy,data = mpg,colour=factor(cyl))
```
task 10:visualize the relationship between engine displacement (displ) and highway miles per gallon (hwy) for cars in the mpg dataset. It creates a scatter plot with both individual data points and a linear regression trend line, where each point is colored according to the number of cylinders
```{r}
qplot(displ,hwy,data = mpg,colour=factor(cyl),geom = c("smooth","point"),method="lm")
```

